Forecasting the old-age dependency ratio to determine the best pension age

### 2019-01-06 PensionAge.Rmd
PensionAge.pdf

---
title: "Forecasting the old age dependency ratio to determine the best pension age"
author:
- familyname: Hyndman
  othernames: Rob J
  address: Department of Econometrics & Business Statistics\newline Monash University, Clayton VIC 3800, Australia
  email: Rob.Hyndman@monash.edu
  correspondingauthor: true
- familyname: Zeng
  othernames: Yijun
  address: Monash University
- familyname: Shang
  othernames: Han Lin
  address: Australian National University
abstract: "We forecast the old-aged dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old aged dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population. We apply these ideas to Australian historical demographic data and the results suggest that the pension age should be increased to 67 by 2021, 70 by 2032 and 75 by 2058 in order to maintain the old aged dependency ratio at the current level, which is 21%. This approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries."
keywords: "functional time series model, coherent forecasts, demographic components, pension and insurance"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: pensionage.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE, cache = TRUE)
# Make sure you have the latest version of rmarkdown and bookdown
# devtools::install_github("rstudio/rmarkdown")
# devtools::install_github("rstudio/bookdown")
# And the latest versions of my packages
# devtools::install_github("robjhyndman/forecast")
# devtools::install_github("robjhyndman/demography")
# devtools::install_github("robjhyndman/addb")
library(demography)
library(addb)
library(ggplot2)
source("functions-v4.R")
```

@@

Australia's ageing population is a result of three factors: (1) significant technological advances in medical care leading to longer life spans; (2) high fertility rates after World War II resulting in the "baby-boomer" generation; and (3) a large decline in fertility rates over the past twenty or so years [@Fehr2008].  Consequently, there are many older people leaving the workforce, and fewer workers replacing them. Similar demographic challenges face many other countries. One way of measuring this problem is via the old-age dependency ratio (OADP): the number of people post-retirement to the number of people of working age [@Walker1982]. The OADP has been increasing for many decades, but recently the rate of increase has jumped markedly, and we show that it will continue to increase at a rapid rate if there are no age-related policy changes.

This is consistent with the Australian Intergenerational Reports [@IGR02; @IGR07; @Swan2010] that explored the extent to which existing age-related policies would affect the burden of future working generations. To date, the Australian government has addressed the problem through a variety of approaches including modifications to compulsory superannuation policies, instigating work incentives for older people, and increasing the pension age (defined as the age at which eligible residents can access the age pension for the first time). In 2009, the federal government announced an increase in pension age from 65 to 67 by 2023 [@Nielson2010]. In May 2014, the Federal Treasurer proposed a further increase in the pension age to 70 by 2035. This new proposal is in line with the recommendations from the @PC13 and the @NCOA2014. In this paper, we investigate the effect of these changes in the pension age on the forecast OADP.

@@

In this paper, we use the old aged dependency ratio as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07].  The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  \text{OADR} = \frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100\%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure.  Forecasts of population structure have been built on forecasts of age and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age and gender-specific population structure.

Using these predicted population paths, we can calculate the average OADR and predict the levels of old age dependency that correspond to different pension age schemes. By testing different pension age schemes, we aim to find the pension age with the desired OADR outcome. Therefore, other contributions of this research are (1) to examine the effect of adjusting pension ages by comparing OADR before and after adjustment; and (2) to suggest a suitable pension age scheme to ensure that the economic burden will remain relatively constant at a desired level in the future. It should be noted that the desired level has been set at 21%, the estimated level of OADR in 2014. Whether this OADR level is the most appropriate level for Australia remains to be examined and is not the focus of this paper. This paper concentrates on providing a framework for policy makers to find a target pension age for a given desired level of OADR.

@@

The Australian government is trying to take early intervention in relation to the pension age to balance between providing social support and fiscal cost. The pension age for males has been set at 65 for over a century. While the pension age for females was 60 years until 1994, it has steadily increased to 65 years, in line with that of males. The previous Labor government announced in 2009 that the pension age will increase six months every two years until 2017, capping at 67 in July 2023. This change is expected to have a significant impact on labor force participation and to decrease the retirement duration of residents because pension age is regarded as a normal age for retirement in Australia. However, the @PC13 proposed further lifting of the pension age to 70 on the grounds that increasing the pension age to 70 would save taxpayers \$150 billion in welfare costs and health spending, and elderly people are capable of continuing work. The National Commission of Audit (NCOA)(2014) holds the same opinion, and recommended lifting the pension age to 70 from 2053. The current Treasurer Joe Hockey announced that the government would increase the pension age to 70 from 2035, which is proposed in the Federal Budget Statement for 2014-2015 to be confirmed by Australian Federal Government. Figure \ref{fig:changespension} illustrates the process of the adjustments on pension age scheme. This announcement received overwhelmingly negative responses in the media. @Guest14 argues from an economic prospective that Australia currently has much better circumstances than other OECD countries in terms of age pension pressure, because spending on the age pension in Australia is  less than half of the OECD average. He also claims that Australia has a higher consumption of all other goods and services, and so why are they not allowed to have more leisure time. Further, @Power14 contends that many Australians are not physically able to work full-time to 70, so that recommendations made by the PC and NCOA are simplistic solutions to a more a complex social problem.

In fact, all supportive arguments for increasing the pension age in Australia have only justified the need to increase the pension age. No justification on the pace or degree of the change has been provided. Also, the government has not disclosed the method it used to determine the new pension age. It is therefore not surprising that the adjustment of the pension age has been questioned by the public. A consistent method to determine the appropriate pension age would be more persuasive in general. For example, it is preferred by OECD countries that pension age should be linked to life expectancy [@OECD12]. There is a practical issue as to how to link life expectancy with pension ages given that there is no information on future mortality rates [@PC13]. @CEDA07 [p7] proposed directly linking pension age with life expectancy in the way that "pension age should increase by approximately 50% of any increase in life expectancy". However, using a "rule of thumb" approach might not be ideal, as 50% of the change in life expectancy may not be a good adjustment to pension age in the long term. My study suggests an approach, linking the pension age with the the aged dependency ratio, based on forecasts of future mortality and fertility rates. My approach has the advantage that it can be used to justify the change in pension age and directly links the pension age with our ultimate goal of a stable financial burden for the public.

@@

In population forecasting, two very different approaches: a deterministic approach (sometimes called a quasi-statistical approach) and a stochastic approach (a purely statistical approach) are frequently used. For instance, the Australian Bureau of Statistics is using a deterministic approach for its demographic projection, while some other countries or institutes forecast demographic trends based on purely statistical approaches such as Statistics New Zealand and the United States Census Bureau. In the deterministic approach, experts project future demographic changes using judgement-based scenarios. It ignores the non-demographic factors such as major pandemics or wars, although historically they are important factors that affect demographic behaviour [@PC13]. In stochastic approaches, the Lee-Carter (LC) method [@LC92] is the most prominent demographic forecasting method, which now has several variants and extensions that relax some of the assumptions of the basic Lee-Carter method. Although statistical forecasting models are rigorous, they still leave a basis for judgment-based adaptation, as they sometimes produce unrealistic projections of demographic factors such as long-run projections of mortality rates [@PC13].

@@

The model proposed by @LC92 is as follows:
\begin{equation}
  \log(m_{x,t})=a_x+b_xk_t+\epsilon_{x,t}
\end{equation}
where $m_{x,t}$ is the central death rate\footnote{The central death rate is properly defined in Section.3} for age $x$ in year $t$. This model involves fitting a matrix of log central rates with sets of age-specific constants, {$a_x$} and {$b_x$}, and a time-varying index $k_t$. $b_x$ is a set of parameters that tell us the rate at which each age responds to changes in $k_t$. $k_t$ measures the general level of the log death rates; $a_x$ is the general pattern across age of the log mortality rate and $\epsilon_{x,t}$ is an error term with zero mean and constant variance, reflecting the randomness that can't be captured by the model.

The model is under-determined in terms of estimation of parameters. For example, if the vector $a,b,k$ is one solution, then for any scalar $c, a-bc, b, k+bc$ also must be one solution. Therefore, Lee and Carter (1992) impose the following constraints:
\begin{equation}
  \sum_{t=1}^{n}k_t=0, \sum_{x=x_1}^{x_m}b_x=1
\end{equation}
Also, we normally seek the least squares solution when estimating parameters. However, in this case, the model cannot be estimated using ordinary regression methods because there is no regressor on the right side of the equation. @LC92 choose a singular value decomposition (SVD) method to undertake PCA, and hence find the least squares solution. It has been shown that applying SVD methods to a matrix of the logarithms of the rates after subtracting the average over time of the log age-specific rates can find the least square solution (Good, 1969). Therefore, the estimations of $b_x$ and $k_t$ are the first principal component and scores of $\log(m_{x,t})-a_x$ (i.e., $b_x=\bm{z_1}$ and $k_t=\bm{u_1}'$  when $\bm{X}=\log(m_{x,t})-a_x$). However, the empirical results suggest that the resulting fitted number of deaths are biased estimates of actual number of deaths because of the log transformation. To overcome this problem, the LC method adjusts $k_t$ such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted $k_t$ is then extrapolated using random walk with drift models, which is equivalent to ARIMA ($0,1,0$) model with $c\neq0$.

@@

@HU07 proposed a nonparametric method for demographic modelling and forecasting. It extended the LC model in four ways:
First, the rates are smoothed using penalized regression splines [@BHT14] before modeling.
Second, it uses functional principal components analysis, which is a continuous version of PCA.
Further, it uses more than one principal component, which address the main weakness of the LC method.
Finally, the forecasting models for the principal component scores are more complex than the Random walk with drift model used in LC model.

The model proposed by @HU07 is as follows:
\begin{align}
  y_t(x) & = s_t(x) + \sigma_t(x)\epsilon_{t,x}\\
  s_t(x) & = \mu(x) + \sum_{j=1}^{J}\beta_{t,j}\phi_j(x) + e_t(x)
\end{align}
The first equation describes the smoothing process. $y_t(x)$ is the observed data in year $t$ and age $x$, and $s_t(x)$ is the smoothed function of $y_t(x)$, estimated by constrained weighted penalized regression splines [BHT14]. $\sigma_t(x)$ allows the variance to be a function of time and age; and $\epsilon_{t,x}$ are independent identical distributed standard normal error terms. The second equation has a similar structure to the LC method, except that it incorporates more than one principal component and score. It describes the dynamics of $s_t(x)$ over time. The $\beta_{t,j}$ for every $j$ and $\mu(x)$ are sets of age-specific constants, and $\phi_j(x)$ for every $j$ is a time-varying index. The $e_t(x)$ is the model error, which is the residual function with mean zero and no serial correlation. The $\mu(x)$ is estimated by the average of $s_t(x)$ across years, which measures the general shape of the observed data $y_t(x)$. The estimates of $\phi_j(x)$ and $\beta_{t,j}$ are the $j$th principal component and scores of $s_t(x)-\mu(x)$ respectively. In other words, we take singular value decomposition of $s_t(x)-\mu(x)$, then equation (7) now becomes
\begin{equation}
  s_t(x)-\mu(x) = \phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_k(x)\beta_{t,k}
\end{equation}
We chose $J$ in a way that we believe the majority of information for $s_t(x)-\mu(x)$ has been captured by the first $J$ components.

In terms of forecasting, by conditioning on the observed data $\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}$ and the set of functional principal components $\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}$, the $h$-step-ahead forecast of log rate $y_{n+h}(x)$ can be acquired by:
\begin{equation}
  \hat{y}_{n+h\mid n}(x)=\hat{s}_{n+h\mid n}(x) = \text{E}[\hat{s}_{n+h\mid n}(x)\mid \bm{I, B}] = \hat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\hat{\beta}_{n+h\mid n,j}
\end{equation}
where $\hat{\beta}_{n+h\mid n,j}$ denotes the $h$-step-ahead forecast of $\beta_{n+h\mid n,j}$. Because of the time-varying property of $\left\{\beta_{t,j}\right\}$, it controls the dynamics of the process. For every component series $\left\{\beta_{t,j}\right\}$, we fit them with an ARIMA$(p_j,d_j,q_j)$ process, where $p_j,d_j,q_j$ are determined by minimizing the Akaike Information Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process as follows:
\begin{equation}
  (1-\phi_1B-\dots-\phi_{p_j}B^{p_j})(1-B)^{d_j}\beta_{t,j}=c+(1+\theta_1B+\dots+\theta_{q_j}B^{q_j})e_t ~\forall j=1,2,\dots,J,
\end{equation}
where the parameters are estimated using maximum likelihood estimation.

@@

We use the same notation and similar procedures to construct data as [@HB08], because the fundamental model for forecasting population for this study is the Hyndman and Ullah method. The data used to construct population age structure included: age and gender-specific birth and death numbers of each calendar year, age and gender-specific population members on 1\ January of each year, and age and gender-specific exposures to risk (i.e., population of age $x$ at 30 June) for each year. The notations are as follows:
\begin{align*}
  B_t(x) &= \text{Births in calendar year $t$ to females of age $x$}\\
  D_t(x) &= \text{Deaths in calendar year $t$ of persons of age $x$}\\
  P_t(x) &= \text{Population of age $x$ at 1 January of year $t$}\\
  E_t(x) &= \text{Population of age $x$ exposed to risk at 30 June of year $t$}\\
\end{align*}
where $x$ denotes age and $x=0,1,2,\dots,m^+$. $m^+$ is the open-ended upper age group, which is 100 in this study. $t$ denotes year and $t=1,2,\dots,n$, where $n$ is the horizon of historical data used. Therefore, all of those data points can be recorded in an $m\times n$ matrix.  The super-scripts $M$ and $F$ denote male and female respectively.

We can obtain historical mortality rates $m_t(x)$ (also called central death rates) and fertility rates $f_t(x)$ as follows:
\begin{align*}
  m_t(x) & = \frac{D_t(x)}{E_t(x)}=\text{age-sex-specific central death rates in calendar year t}\\
  f_t(x) & = B_t(x)/E^F_t(x)=\text{age-specific fertility rates in calendar year t}
\end{align*}

In addition, the net migration\footnote{Net migration is the difference between immigration and emigration in a certain area during a specified time frame.} (denoted $G$) is estimated using a demographic growth-balance equation, which is an equation to summarize the relationship between population change and three factors: net migration, births and deaths. The equation is:
\begin{align*}
  G_t(x,x+1)     & =P_{t+1}(x+1)-P_t(x)+D_t(x,x+1)\text{~for~} x=0,1,2,\dots,p-2,\\
  G_t(p-1^+,p^+) & =P_{t+1}(p^+)-P_t(p^+)-P_t(p-1)+D_t(p-1^+,p^+)\\
  G_t(B,0)       & =P_{t+1}(0)-B_t+D_t(B,0)
\end{align*}
where $G_t(x,x+1)$ refers to net migration in calendar year $t$ of persons aged $x$ at beginning of year $t$, $G_t(p-1^+,p)$ refers to migration in calendar year $t$ of persons aged $p-1$ and older at beginning of year $t$, and $G_t(B,0)$ refers to deaths in calendar year $t$ of births during year $t$; and similarly for deaths, $D_t(x,x+1), D_t(p-1^+,p^+), D_t(B,0)$. The deaths are estimated using the standard life table approach of population projection. It should be noted that the estimated net migration includes errors in data recording.

Historical central death rates and start-year and mid-year populations of residents in Australia by sex and age in single years for age group 0-99 and 100+ have been sourced from the @HMD for 1921--2009. The start-year population for 2010 is also used to estimate net migration. Data for age-specific annual fertility rates by single years of age for 15-49 over the period 1950-2009 was obtained from the @ABS12. Even though annual mortality rates are available from 1921 to 2009, only data collected after 1950 is used for two reasons. First, the period between 1921-1950 has very different mortality patterns from later years because of wars and epidemics [@HB08]. Since the model for this study is valid under the assumption that extrapolative methods can be used, it seemed best to delete the less relevant data, which might influence estimations of parameters if we include them. Secondly, fertility rates are only available from 1950, so the mortality rate was cut to fit the same period.

@@

```{r pension, dependson="OADRat65"}
### find the optimal scheme of pension.age based on mean forecast
target <- oadp.f$mean[4]

>>>>

---
title: "Forecasting the old-age dependency ratio to determine the best pension age"
author:
- familyname: Hyndman
  othernames: Rob J
  address: Department of Econometrics & Business Statistics\newline Monash University, Clayton VIC 3800, Australia
  email: Rob.Hyndman@monash.edu
  correspondingauthor: true
- familyname: Zeng
  othernames: Yijun
  address: Monash University
- familyname: Shang
  othernames: Han Lin
  address: Australian National University
abstract: "We forecast the old-age dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old-age dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population. We apply these ideas to Australian historical demographic data and the results suggest that the pension age should be increased to 67 by 2021, 70 by 2032 and 75 by 2058 in order to maintain the old-age dependency ratio at the current level, which is 21%. This approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries."
keywords: "functional time series model, coherent forecasts, demographic components, pension and insurance"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: pensionage.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE, cache = TRUE)
# Make sure you have the latest version of rmarkdown and bookdown
# devtools::install_github("rstudio/rmarkdown")
# devtools::install_github("rstudio/bookdown")
# And the latest versions of Our packages
# devtools::install_github("robjhyndman/forecast")
# devtools::install_github("robjhyndman/demography")
# devtools::install_github("robjhyndman/addb")
library(demography)
library(addb)
library(ggplot2)
source("functions-v4.R")
```

@@

Australia's ageing population is a result of three factors: (1) significant technological advances in medical care leading to longer life spans; (2) high fertility rates after World War II resulting in the "baby-boomer" generation; and (3) a large decline in fertility rates over the past twenty or so years [@Fehr2008].  Consequently, there are many older people leaving the workforce, and fewer workers replacing them. Similar demographic challenges face many other countries. One way of measuring this problem is via the old-age dependency ratio (OADR): the number of people post-retirement to the number of people of working age [@Walker1982]. The OADR has been increasing for many decades, but recently the rate of increase has jumped markedly, and we show that it will continue to increase at a rapid rate if there are no age-related policy changes.

This is consistent with the Australian Intergenerational Reports [@IGR02; @IGR07; @Swan2010] that explored the extent to which existing age-related policies would affect the burden of future working generations. To date, the Australian government has addressed the problem through a variety of approaches including modifications to compulsory superannuation policies, instigating work incentives for older people, and increasing the pension age (defined as the age at which eligible residents can access the age pension for the first time). In 2009, the federal government announced an increase in pension age from 65 to 67 by 2023 [@Nielson2010]. In May 2014, the Federal Treasurer proposed a further increase in the pension age to 70 by 2035. This new proposal is in line with the recommendations from the @PC13 and the @NCOA2014. In this paper, we investigate the effect of these changes in the pension age on the forecast OADR.

@@

In this paper, we use the OADR as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07].  The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  \text{OADR} = \frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100\%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure.  Forecasts of population structure have been built on forecasts of age- and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age- and gender-specific population structure.

Using these predicted population paths, we can calculate the average OADR and predict the levels of old-age dependency that correspond to different pension age schemes. By testing different pension age schemes, we aim to find the pension age with the desired OADR outcome. Therefore, other contributions of this research are (1) to examine the effect of adjusting pension ages by comparing OADR before and after adjustment; and (2) to suggest a suitable pension age scheme to ensure that the economic burden will remain relatively constant at a desired level in the future. It should be noted that the desired level has been set at 21%, the estimated level of OADR in 2014. Whether this OADR level is the most appropriate level for Australia remains to be examined and is not the focus of this paper. This paper concentrates on providing a framework for policy makers to find a target pension age for a given desired level of OADR.

@@

The Australian government is trying to take early intervention in relation to the pension age to balance between providing social support and fiscal cost. The pension age for males has been set at 65 for over a century. While the pension age for females was 60 years until 1994, it has steadily increased to 65 years, in line with that of males. The previous Labor government announced in 2009 that the pension age will increase six months every two years until 2017, capping at 67 in July 2023. This change is expected to have a significant impact on labor force participation and to decrease the retirement duration of residents because pension age is regarded as a normal age for retirement in Australia. However, the @PC13 proposed further lifting of the pension age to 70 on the grounds that increasing the pension age to 70 would save taxpayers \$150 billion in welfare costs and health spending, and elderly people are capable of continuing work. The National Commission of Audit (NCOA) (2014) holds the same opinion, and recommended lifting the pension age to 70 from 2053. The current Treasurer Joe Hockey announced that the government would increase the pension age to 70 from 2035, which is proposed in the Federal Budget Statement for 2014-2015 to be confirmed by Australian Federal Government. Figure \ref{fig:changespension} illustrates the process of the adjustments on pension age scheme. This announcement received overwhelmingly negative responses in the media. @Guest14 argues from an economic prospective that Australia currently has much better circumstances than other OECD countries in terms of age pension pressure, because spending on the age pension in Australia is  less than half of the OECD average. He also claims that Australia has a higher consumption of all other goods and services, and so why are they not allowed to have more leisure time. Further, @Power14 contends that many Australians are not physically able to work full-time to 70, so that recommendations made by the PC and NCOA are simplistic solutions to a more a complex social problem.

In fact, all supportive arguments for increasing the pension age in Australia have only justified the need to increase the pension age. No justification on the pace or degree of the change has been provided. Also, the government has not disclosed the method it used to determine the new pension age. It is therefore not surprising that the adjustment of the pension age has been questioned by the public. A consistent method to determine the appropriate pension age would be more persuasive in general. For example, it is preferred by OECD countries that pension age should be linked to life expectancy [@OECD12]. There is a practical issue as to how to link life expectancy with pension ages given that there is no information on future mortality rates [@PC13]. @CEDA07 [p7] proposed directly linking pension age with life expectancy in the way that "pension age should increase by approximately 50% of any increase in life expectancy". However, using a "rule of thumb" approach might not be ideal, as 50% of the change in life expectancy may not be a good adjustment to pension age in the long term. Our study suggests an approach, linking the pension age with the the aged dependency ratio, based on forecasts of future mortality and fertility rates. Our approach has the advantage that it can be used to justify the change in pension age and directly links the pension age with our ultimate goal of a stable financial burden for the public.

@@

In population forecasting, two very different approaches: a deterministic approach (sometimes called a quasi-statistical approach) and a stochastic approach (a purely statistical approach) are frequently used. For instance, the Australian Bureau of Statistics (ABS) is using a deterministic approach for its demographic projection, while some other countries or institutes forecast demographic trends based on purely statistical approaches such as Statistics New Zealand and the United States Census Bureau. In the deterministic approach, experts project future demographic changes using judgement-based scenarios. It ignores the non-demographic factors such as major pandemics or wars, although historically they are important factors that affect demographic behaviour [@PC13]. In stochastic approaches, the Lee-Carter (LC) method [@LC92] is the most prominent demographic forecasting method, which now has several variants and extensions that relax some of the assumptions of the basic Lee-Carter method [see, e.g., @SBH11]. Although statistical forecasting models are rigorous, they still leave a basis for judgment-based adaptation, as they sometimes produce unrealistic projections of demographic factors such as long-run projections of mortality rates [@PC13].

@@

The model proposed by @LC92 is as follows:
\begin{equation}
  \ln(m_{x,t})=a_x+b_xk_t+\epsilon_{x,t}
\end{equation}
where $m_{x,t}$ is the central death rate\footnote{The central death rate is properly defined in Section~\ref{sec:data}} for age $x$ in year $t$. This model involves fitting a matrix of log central rates with sets of age-specific constants, {$a_x$} and {$b_x$}, and a time-varying index $k_t$. $b_x$ is a set of parameters that tell us the rate at which each age responds to changes in $k_t$. $k_t$ measures the general level of the log death rates; $a_x$ is the general pattern across age of the log mortality rate and $\epsilon_{x,t}$ is an error term with zero mean and constant variance, reflecting the randomness that can't be captured by the model.

The model is under-determined in terms of estimation of parameters. For example, if the vector $a,b,k$ is one solution, then for any scalar $c, a-bc, b, k+bc$ also must be one solution. Therefore, Lee and Carter (1992) impose the following constraints:
\begin{equation}
  \sum_{t=1}^{n}k_t=0, \sum_{x=x_1}^{x_m}b_x=1
\end{equation}
Also, we normally seek the least squares solution when estimating parameters. However, in this case, the model cannot be estimated using simple regression methods because there is no regressor on the right side of the equation. @LC92 choose a singular value decomposition (SVD) method to undertake principal component analysis (PCA), and hence find the least squares solution. It has been shown that applying SVD method to a matrix of the logarithms of the rates after subtracting the average over time of the log age-specific rates can find the least square solution (Good, 1969). Therefore, the estimations of $b_x$ and $k_t$ are the first principal component and scores of $\ln(m_{x,t})-a_x$ (i.e., $b_x=\bm{z_1}$ and $k_t=\bm{u_1}'$  when $\bm{X}=\ln(m_{x,t})-a_x$). However, the empirical results suggest that the resulting fitted number of deaths are biased estimates of actual number of deaths because of the $\ln$ transformation. To overcome this problem, the LC method adjusts $k_t$ such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted $k_t$ is then extrapolated using random walk with drift models, which is equivalent to ARIMA ($0,1,0$) model with $c\neq0$.

@@

@HU07 proposed a nonparametric method for demographic modelling and forecasting. It extended the LC model in four ways:
\begin{enumerate}
\item The rates are smoothed using penalized regression splines [@BHT14] before modeling.
\item It uses functional principal components analysis, which is a continuous version of PCA.
\item It uses more than one principal component, which address the main weakness of the LC method.
\item The forecasting models for the principal component scores are more complex than the Random walk with drift model used in LC model.
\end{enumerate}

The model proposed by @HU07 is as follows:
\begin{align}
  y_t(x) & = s_t(x) + \sigma_t(x)\epsilon_{t,x}\label{eq:HU_1}\\
  s_t(x) & = \mu(x) + \sum_{j=1}^{J}\beta_{t,j}\phi_j(x) + e_t(x)\label{eq:HU_2}
\end{align}
Equation \eqref{eq:HU_1} describes the smoothing process. $y_t(x)$ is the observed data in year $t$ and age $x$, and $s_t(x)$ is the smoothed function of $y_t(x)$, estimated by constrained weighted penalized regression splines [@BHT14]. $\sigma_t(x)$ allows the variance to be a function of time and age; and $\epsilon_{t,x}$ are independent identical distributed standard normal error terms. Equation \eqref{eq:HU_2} has a similar structure to the LC method, except that it incorporates more than one principal component and score. It describes the dynamics of $s_t(x)$ over time. The $\beta_{t,j}$ for every $j$ and $\mu(x)$ are sets of age-specific constants, and $\phi_j(x)$ for every $j$ is a time-varying index. The $e_t(x)$ is the model error, which is the residual function with mean zero and no serial correlation. The $\mu(x)$ is estimated by the average of $s_t(x)$ across years, which measures the general shape of the observed data $y_t(x)$. The estimates of $\phi_j(x)$ and $\beta_{t,j}$ are the $j$th principal component and scores of $s_t(x)-\mu(x)$ respectively. In other words, we take singular value decomposition of $s_t(x)-\mu(x)$, then \eqref{eq:HU_2} now becomes
\begin{equation*}
  s_t(x)-\mu(x) = \phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_k(x)\beta_{t,k}
\end{equation*}
We chose $J$ in a way that we believe the majority of information for $s_t(x)-\mu(x)$ has been captured by the first $J$ components.

In terms of forecasting, by conditioning on the observed data $\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}$ and the set of functional principal components $\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}$, the $h$-step-ahead forecast of log rate $y_{n+h}(x)$ can be acquired by:
\begin{equation}
  \widehat{y}_{n+h\mid n}(x)=\widehat{s}_{n+h\mid n}(x) = \text{E}[\widehat{s}_{n+h\mid n}(x)\mid \bm{I, B}] = \widehat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\widehat{\beta}_{n+h\mid n,j}
\end{equation}
where $\widehat{\beta}_{n+h\mid n,j}$ denotes the $h$-step-ahead forecast of $\beta_{n+h\mid n,j}$. Because of the time-varying property of $\left\{\beta_{t,j}\right\}$, it controls the dynamics of the process. For every component series $\left\{\beta_{t,j}\right\}$, we fit them with an ARIMA$(p_j,d_j,q_j)$ process, where $p_j,d_j,q_j$ are determined by minimizing the Akaike Information Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process is given as follows:
\begin{equation*}
  (1-\phi_1B-\dots-\phi_{p_j}B^{p_j})(1-B)^{d_j}\beta_{t,j}=c+(1+\theta_1B+\dots+\theta_{q_j}B^{q_j})e_t, \quad ~\forall j=1,2,\dots,J,
\end{equation*}
where the parameters are estimated using maximum likelihood estimation.

@@

We use the same notation and similar procedures to construct data as in @HB08, because the fundamental model for forecasting population for this study is the @HU07 method. The data used to construct population age structure included: age- and gender-specific birth and death numbers of each calendar year, age- and gender-specific population members on 1\ January of each year, and age- and gender-specific exposures to risk (i.e., population of age $x$ at 30 June) for each year. The notations are as follows:
\begin{align*}
  B_t(x) &= \text{Births in calendar year $t$ to females of age $x$}\\
  D_t(x) &= \text{Deaths in calendar year $t$ of persons of age $x$}\\
  P_t(x) &= \text{Population of age $x$ at 1 January of year $t$}\\
  E_t(x) &= \text{Population of age $x$ exposed to risk at 30 June of year $t$}\\
\end{align*}
where $x$ denotes age and $x=0,1,2,\dots,m^+$. $m^+$ is the open-ended upper age group, which is 100 in this study. $t$ denotes year and $t=1,2,\dots,n$, where $n$ is the horizon of historical data used. Therefore, all of those data points can be recorded in an $m\times n$ matrix.  The super-scripts $M$ and $F$ denote male and female respectively.

We can obtain historical mortality rates $m_t(x)$ (also called central death rates) and fertility rates $f_t(x)$ as follows:
\begin{align*}
  m_t(x) & = \frac{D_t(x)}{E_t(x)}=\text{age-sex-specific central death rates in calendar year t}\\
  f_t(x) & = \frac{B_t(x)}{E^F_t(x)}=\text{age-specific fertility rates in calendar year t}
\end{align*}

In addition, the net migration\footnote{Net migration is the difference between immigration and emigration in a certain area during a specified time frame} (denoted $G$) is estimated using a demographic growth-balance equation, which is an equation to summarize the relationship between population change and three factors: net migration, births and deaths. The equation is expressed as: \todo{Please define $p$ and $^+$}
\begin{align*}
  G_t(x,x+1)     & =P_{t+1}(x+1)-P_t(x)+D_t(x,x+1)\text{~for~} x=0,1,2,\dots,p-2,\\
  G_t(p-1^+,p^+) & =P_{t+1}(p^+)-P_t(p^+)-P_t(p-1)+D_t(p-1^+,p^+)\\
  G_t(B,0)       & =P_{t+1}(0)-B_t+D_t(B,0)
\end{align*}
where $G_t(x,x+1)$ refers to net migration in calendar year $t$ of persons aged $x$ at beginning of year $t$, $G_t(p-1^+,p)$ refers to migration in calendar year $t$ of persons aged $p-1$ and older at beginning of year $t$, and $G_t(B,0)$ refers to deaths in calendar year $t$ of births during year $t$; and similarly for deaths, $D_t(x,x+1), D_t(p-1^+,p^+), D_t(B,0)$. The deaths are estimated using the standard life table approach of population projection. It should be noted that the estimated net migration includes errors in data recording.

Historical central death rates and start-year and mid-year populations of residents in Australia by sex and age in single years for age group 0-99 and 100+ have been sourced from the @HMD for 1921--2014. The start-year population for 2015 is also used to estimate net migration. Data for age-specific annual fertility rates by single years of age for 15-49 over the period 1950-2014 was obtained from the @ABS12. Even though annual mortality rates are available from 1921 to 2009, only data collected after 1950 is used for two reasons. First, the period between 1921-1950 has very different mortality patterns from later years because of wars and epidemics [@HB08]. Since the model for this study is valid under the assumption that extrapolative methods can be used, it seemed best to delete the less relevant data, which might influence estimations of parameters if we include them. Secondly, fertility rates are only available from 1950, so the mortality rate was cut to fit the same period. \todo{Han has updated the data period.}

@@

```{r pension, dependson="OADRat65"}
### find the optimal scheme of pension.age based on mean forecast
target <- OADR.f$mean[4]
